---
title: "Survey quantiles with dummy sample"
author: "Sara Hänzi"
date: today
date-format: "D MMM YYYY"
format: gfm
---

# Preparation

package loads 
```{r warning=FALSE, message=FALSE}
library(survey)
library(dplyr)
library(ggplot2)
library(patchwork)
```

check survey version
```{r}
packageVersion("survey")
```


define quantiles to be estimated, some not-quite-straightforward numbers as some examples only pop up with special cases
```{r}
quantiles <- c(0.01, 0.05, 0.1, 0.15, seq(21, 81, 10)*0.01, 0.85, 0.9, 0.95, 0.99)
```

source all functions

```{r}
source("functions_for_estimation_and_plotting.R")
```

# No stratification 

dummy population with 
N = n = 10

## prepare data and design
```{r}
data_no_strat <- data.frame(x = 1:10, id = 1:10, fpc = 10, probs = 1)

design_no_strat <- svydesign(id = ~id, 
                             probs = ~probs, 
                             data = data_no_strat, 
                             fpc = ~fpc)
```

## example call
```{r}
res <- svyquantile(~x, design_no_strat, quantiles, ci = TRUE,
            interval.type = "mean", qrule = "hf1")
```

Plot the estimated quantiles (the color indicates how many NAs are present in the confidence limits). The confidence interval is plotted in shaded gray as an area, however this is not visible when the interval is the same as the quantile estimate (which is the expected behaviour here as everything is known).
```{r}
quant_df <- quant_to_df(res, quantiles)
plot_quantile_ci(quant_df, "mean and hf1")
```

get the proportion that for which it is true that x >= the estimated quantile; checking this as this seems to be used in the implementation for the estimation of the confidence intervals
```{r}
a <- purrr::map(quant_df$quantile, .f = \(z) survey::svymean(~ x <= z, design_no_strat)) %>% 
  purrr::map_dbl(2)

plot_prop_quantile(quant_df$level, a)
```


## loop over different combinations
do for different versions / combinations of `interval_type` and `qrule`

### with newsvyquantile
```{r}
quantile_plot_loop(design_no_strat, quantiles, 
                   qrules = c("hf1", "hf7", "hf8"),
                   interval_types = c("mean", "beta", "xlogit"))
```

### with oldsvyquantile

```{r}
old_quantile_plot_loop(design_no_strat, quantiles,
                       ties = c("discrete", "rounded"),
                       interval_types = c("Wald", "betaWald"))
```


# dummy population with stratification

overall: n = 20, N = 110

stratum 1: n = 10, N = 10, everything is known

stratum 2: n = 10, N = 100

```{r}
data_strat <- data.frame(x = c(1:10, 1:10), id = 1:20, 
                         strata =  c(rep(1, 10), rep(2,10)),
                         fpc = c(rep(10, 10), rep(100, 10)), 
                         probs = c(rep(1, 10),rep(0.1,10)))

design_strat <- svydesign(id = ~id, 
                          strata = ~strata,
                          probs = ~probs,
                          data = data_strat, 
                          fpc = ~fpc)
```

## example call
```{r}
res <- svyquantile(~x, design_strat, quantiles, ci = TRUE,
            interval.type = "mean", qrule = "hf1")

res
```

Plot the quantiles and their confidence intervals. Again, the color indicates how many NAs are present in the confidence limits and the confidence interval is plotted in shaded gray as an area.
```{r}
quant_df <- quant_to_df(res, quantiles)
plot_quantile_ci(quant_df, "mean and hf1")
```
### loop with new svyquantile
```{r}
quantile_plot_loop(design_strat, quantiles,
                   qrules = c("hf1", "hf7", "hf8"),
                   interval_types = c("mean", "beta", "xlogit"))
```

### loop with old svyquantile

```{r}
old_quantile_plot_loop(design_strat, quantiles,
                       ties = c("discrete", "rounded"),
                       interval_types = c("Wald", "betaWald"))
```

